Reliable feature point matching without the use of geometric constraints

### 2013-06-06 text/clustering.tex
V. 2.0

\documentclass[12pt,journal,compsoc]{IEEEtran}
%\documentclass{article}
\usepackage{footnote}
\usepackage{sidecap}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{hyperref}
\usepackage[stable]{footmisc}
\usepackage{caption}
\usepackage{adjustbox}
\usepackage{subcaption}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{amsmath}
\usepackage{amssymb}

@@

angles. From these pairs we can create test suites by cropping out $250$ 
pixel by $250$ pixel patches of the original pair with a random vertical 
and horizontal offset. Given a source image pair we can obtain $n$ pairs 
of patches where the two patches in any pair might or might not overlap.  
This ensures that the patch pairs whith no overlap still retain a basic 
similarity to each other, while the patches that do overlap often only 
overlap on a small part of their surface. Figure \ref{fig:fairey} shows 
an example of possible pairs of test images (\ref{fig:fairey_crop}) 
produced from two source images (\ref{fig:fairey1} and 
\ref{fig:fairey2}). Producing $n$ such pairs allows us to test not just 
how well the matching algorithm performs on corresponding images, but 
also how many false positives we get on similar images that don't 
overlap.\footnote{The set of source images and homographies as well as 
	the a script to generate the cropped test sets based on them have 
	been made available online together with the exact test sets used in 
this paper on INSERT URL HERE}.

@@

%	\makebox[0.5\textwidth][c]{%
%		\begin{subfigure}[t]{0.12\textwidth}
%			\includegraphics[width=\textwidth]{images/pair_example1}
%			\caption{Mural by Fairey (1)}
%			\label{fig:fairey1}
%		\qquad %add desired spacing between images, e. g. ~, \quad, 
%		\begin{subfigure}[t]{0.12\textwidth}
%			\includegraphics[width=\textwidth]{images/pair_example2}
%			\caption{Mural by Fairey (2)}
%			\label{fig:fairey2}
%	\\
	%\makebox[0.5\textwidth][c]{
	%	%~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
	%	\begin{subfigure}[t]{0.5\textwidth}
	%}%

@@

	\includegraphics[width=0.5\textwidth]{images/crop_examples}

@@

\subsection{Experimental setup}

@@

algorithms have been tested against a the algorithm Lowe proposed in 
\cite{lowe2004sift} as well as the isodata matching algorithm proposed 
in \cite{das2008event}. For each test set of cropped image pairs the 
four algorithms\footnote{\emph{MM}, \emph{MMC}, \emph{SIFT} and 
\emph{Isodata}} are applied to each pair of images ($100$ per test set), 
and the threshold for each algorithm is varied over 50 evenly spaced 
values.  For \emph{MM} and \emph{SIFT} these values lie between $0.4$ 
and $0.8$ where as the threshold for \emph{MMC} and \emph{Isodata} is 
varied between $0.8$ and $1.0$.  For each match returned by any of the 
algorithms, the match is registered as an inlier if the endpoint in the 
second image lies within $5$px of the true position calculated by the 
homography.  That is, given $H$ relating $I_1$ with $I_2$ and a match 
($p_1, p_2$) then the match is an inlier if $\left\vert Hp_1 - p_2 
\right\vert + \left\vert H^{-1}p_2 - p_1 \right\vert < 5$. The 

@@

@@

			\includegraphics[width=\textwidth]{images/scharf}
			\caption{Faces (Scharf)}
			\label{fig:result_faces}

@@

			\includegraphics[width=\textwidth]{images/result_scharf}

@@

@@

Figure \ref{fig:accumulated} shows the results on the test set proposed 
earlier\footnote{Illustrated in Figure \ref{mural}}. The performance of 
the proposed algorithm is shown to be [FILL IN ONCE THE DATA IS READY] 
to \emph{SIFT} and \emph{Isodata} when testing on image pairs that might 
not overlap. This result is repeated (TODO: is it?) when testing on the 
graf image as demonstrated in Figure \ref{fig:result_graf}. To 
illustrate that this performance improvement isn't limited to street 
art, figure \ref{fig:gallagher} shows the performance on an image from 
the Gallagher data set \cite{gallagher2008} where images featuring 
occlusion and change of perspective are matched. While performance is of 
course affected by the magnitude of for example view port change, figure 
\ref{fig:result_viewport} shows that \emph{MMC} consistently outperforms 
the \emph{SIFT} algorithm over different magnitudes of change in 
perspective. TODO: include note on plot for different descriptors.

In terms of performance, table \ref{table:running_times} shows the 
running time of the four algorithms over $100$ image pairs of $250$px by 
$250$px. These numbers should be interpreted with the knowledge that 
much of the code behind \emph{MMC} and \emph{Isodata} is implemented in 
python while \emph{MM} and \emph{SIFT} makes use of opencv to execute 
computationally intensive operations in C++ which renders it much 
faster.
%
\begin{savenotes}
\begin{table}
	\centering
	\small
\begin{tabular}{l*{4}{c}}
	Algorithm: & \emph{SIFT} & \emph{MM} & \emph{MMC} & \emph{Isodata} 
	\\
	\noalign{\smallskip} 
	%
	Running time: & 20.94s & 18.23s & 883.99s & 555.33s \\
\end{tabular}
\caption{Running times\footnote{As tested on a Intel(R) Core(TM) i5-3550 
CPU @ 3.30GHz with 8Gb memory}}
\label{table:running_times}
\end{table}
\end{savenotes}
%
\section{Summary}
We have addressed the problem of matching feature points without using 
geometrical constraints and proposed two algorithms \emph{Mirror Match 
(MM)} and \emph{Mirror Match with Clustering (MMC)}. The two algorithms 
share the same common idea that feature points should have better 
matches in another image than in the image they came from to be 
considered good matches and \emph{MMC} further improves on this idea by 
using the structure of the similarity graph of the feature points. We 
also address how to test the algorithms by constructing a image set of 
$8$ source image pairs and a way to generate testsets based on these 
pairs. The proposed approach shows promising results when comparing the 
algorithms on $9$ testsets of $100$ image pairs each where many of the 
pairs might not have any overlapping areas. In future work it will be 
usefull to investigate how the algorithms perform when testing on sets 
of images instead of pairs.
%
\bibliographystyle{IEEEtran}
\bibliography{bibliography}
\end{document}

>>>>

\documentclass[12pt,journal,compsoc]{IEEEtran}
%\documentclass{article}
\usepackage{footnote}
\usepackage[font={small}]{caption}
\usepackage{sidecap}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{hyperref}
\usepackage[stable]{footmisc}
\usepackage{caption}
\usepackage{adjustbox}
\usepackage{subcaption}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{amsmath}
\usepackage{amssymb}

@@

angles.  From these pairs we can create test suites by cropping out 
$250$ pixel by $250$ pixel patches of the original pair with a random 
vertical and horizontal offset. Given a source image pair we can obtain 
$n$ pairs of patches where the two patches in any pair might or might 
not overlap.  This ensures that the patch pairs whith no overlap still 
retain a basic similarity to each other, while the patches that do 
overlap often only overlap on a small part of their surface. This 
dataset will from here on be referred to as the \emph{Murals} dataset.  
Figure \ref{fig:fairey} shows an example of possible pairs of test 
images produced from two source images to the left.  Producing $n$ such 
pairs allows us to test not just how well the matching algorithm 
performs on corresponding images, but also how many false positives we 
get on similar images that don't overlap.\footnote{The set of source 
	images and homographies as well as the script to generate the 
cropped test sets based on them have been made available online together 
with the exact test sets used in this paper on 
\href{https://github.com/arnfred/Murals}{www.github.com/arnfred/Murals}}.

@@

		\begin{subfigure}[t]{0.044\textwidth}

@@

			\includegraphics[width=\textwidth]{images/pair_example1}
			\label{fig:fairey1}
		~ %add desired spacing between images, e. g. ~, \quad, (or a 
		%blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{0.044\textwidth}
			\includegraphics[width=\textwidth]{images/pair_example2}
			\label{fig:fairey2}
		~ %
		\begin{subfigure}[t]{0.37\textwidth}
			\includegraphics[width=\textwidth]{images/crop_examples}

@@

%\subsection{Experimental setup}

@@

algorithms have been tested against \emph{SIFT}, the algorithm Lowe 
proposed in \cite{lowe2004sift} as well as the \emph{Isodata} matching 
algorithm proposed in \cite{das2008event} which uses geometric 
information to improve the matching. For each test set of cropped image 
pairs the four algorithms\footnote{\emph{MM}, \emph{MMC}, \emph{SIFT} 
and \emph{Isodata}} are applied to each pair of images ($100$ per test 
set), and the threshold for each algorithm is varied over 50 evenly 
spaced values. For each match returned by any of the algorithms, the 
match is registered as an inlier if given $H$ relating $I_1$ with $I_2$ 
and a match ($p_1, p_2$) the match is an inlier if $\left\vert Hp_1 - 
p_2 \right\vert + \left\vert H^{-1}p_2 - p_1 \right\vert < 5$. The 

@@

	\makebox[0.5\textwidth][c]{%
		\begin{subfigure}[t]{.15\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/scharf}
			\caption{Faces (Scharf)}
			\label{fig:result_faces}
		\end{subfigure}%
		~ %add desired spacing between images, e. g. ~, \quad, \qquad		  
		%(or a blank line to force the subfigure onto a new line)
		\begin{subfigure}[t]{.27\textwidth}
			\centering
			\includegraphics[width=\textwidth]{images/result_scharf}
			%\caption{Performance on Scharf}
		\end{subfigure}%
	}%

@@

			\includegraphics[width=\textwidth]{images/pitts}
			\caption{Pittsburg \cite{gallagher2008}}
			\label{fig:result_pitts}

@@

			\includegraphics[width=\textwidth]{images/result_pitts}

@@

	\includegraphics[width=0.45\textwidth]{images/result_accumulated}
	\caption{Results on the mural dataset. For each of the 8 source 
		image pairs, 100 test pairs have been created. The x-axis 
		illustrates the accumulated returned matches for all pairs.
		changes.}
	\label{fig:result_accumulated}

@@

Figure \ref{fig:result_accumulated} shows the accumulated results on the 
\emph{Murals} dataset\footnote{Illustrated in Figure \ref{mural}}. It 
shows that \emph{MMC} and \emph{MM} is overall superior to both 
\emph{SIFT} and \emph{Isodata} which uses geometric constraints when 
tested on image pairs with no assurance of overlap depicting simple 
planar scenes. Figure \ref{fig:result_faces} illustrates the result on a 
particular image pair from the \emph{Murals} dataset. For an example of 
a real life use case \ref{fig:result_pitts} shows the results on 
potentially overlapping image pairs generated from a typical holiday 
scene with partial overlap and slight view port change. As can be seen 
in figure \ref{fig:result_viewport}, \emph{MMC} consistently outperforms 
\emph{SIFT} across different magnitudes of the viewport change. For the 
results on image 1 and 3 from the \emph{Graf} dataset, see figure 
\ref{fig:result_graf}.
Generally all plots show good results by \emph{Isodata} when only the 
best matches are returned, but when the thresholds are more lenient the 
quality of the returned matches degrades quickly and \emph{MMC} and 
\emph{MM} yield better results.

In terms of performance, table \ref{table:running_times} shows the 
running time of the four algorithms over $100$ image pairs of $250$px by 
$250$px. These numbers should be interpreted with the knowledge that 
much of the code behind \emph{MMC} and \emph{Isodata} is implemented in 
python while \emph{MM} and \emph{SIFT} makes use of opencv to execute 
computationally intensive operations in C++ which renders it much 
faster.
%
\begin{savenotes}
\begin{table}
	\centering
	\small
\begin{tabular}{l*{4}{c}}
	Algorithm: & \emph{SIFT} & \emph{MM} & \emph{MMC} & \emph{Isodata} 
	\\
	\noalign{\smallskip} 
	%
	Running time: & 20.94s & 18.23s & 883.99s & 555.33s \\
\end{tabular}
\caption{Running times\footnote{As tested on a Intel(R) Core(TM) i5-3550 
CPU @ 3.30GHz with 8Gb memory}}
\label{table:running_times}
\end{table}
\end{savenotes}
%
\section{Summary}
We have addressed the problem of matching feature points without using 
geometrical constraints and proposed two algorithms \emph{Mirror Match 
(MM)} and \emph{Mirror Match with Clustering (MMC)}. The two algorithms 
share the same common idea that feature points should have better 
matches in another image than in the image they came from to be 
considered good matches and \emph{MMC} further improves on this idea by 
using the structure of the similarity graph of the feature points. We 
also address how to test the algorithms by constructing the 
\emph{Murals} dataset consisting of $8$ source image pairs and a way to 
generate testsets based on these pairs.  The proposed approach shows 
promising results on both the \emph{Murals} dataset as well as a real 
life example from the Gallagher dataset and the \emph{Graf} image set, 
beating even a method using geometrical constraints when matching image 
pairs that might not overlap.
%
\bibliographystyle{IEEEtran}
\bibliography{bibliography}
\end{document}
